Considering scale in multilayer animal social networks

### 2020-06-15 paper/manuscript.Rmd
big discuss

# Discussion
<!-- TODO: edge overlap can indicate if some associations are dependent on context -->
<!-- sufficient observations is essential for capturing reliable networ -->

@@

<!-- TODO add? each ordinal network layer consisted of avg 380 fixes * ID -->

@@

## Phenotypes and Individual fitness

@@

For instance, social scale is major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues unlike in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

Temporal scale is also important for understanding the repercussions of social
phenotypes. Varying the time window length and position for caribou, we found
there was a lot of variation in the multilayer network metrics (Figure
\@ref(fig:winlen), Figure \@ref(fig:winpos)). These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest. For instance, social phenotypes can change over ontogeny
in ways that can affect fitness [@Berger_2015; @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can make defining appropriate temporal
scales for testing the fitness implications of an individual’s social phenotype
over different stages of ontogeny even more difficult when they vary by
individual. However, considering fine temporal scales provide insight across
developmental stages that may otherwise be missed [@Turner_2017]. Defining
temporal scales is a challenge in multilayer networks just as in monolayer
networks; however, multilayer networks provide metrics that enable researchers
to look at similarities and differences in separate layers representing
different stages of ontogeny, or seasons as in our case study, which can be
difficult to detect.

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes. The multilayer network framework enables
researchers to examine these different social scales simultaneously rather than
individually, which is what has been done under most circumstances, but is not
very realistic. Metrics from multilayer networks provide an opportunity to
simplify statistical models. Interactions that happen side-by-side, proximity
associations, and long-distance interactions can be encapsulated in a single
multilayer network, and a single metric can be used to describe the various
social scales of an individual’s network position. This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful reduction of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014]. Furthermore, multilayer networks provide a framework
to test hypotheses linking social phenotypes and fitness across temporal scales.
Because of the non-independent nature of social interactions and associations,
especially over time, it is hard to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

@@

Although multilayer animal social networks are relatively novel [@Silk_2018],
movement behaviour within multilayer networks has been considered. Spatial
networks are physical locations (nodes) which are connected in a network when
individuals move between said locations (edges). For example, spatial networks
were applied to monitor movement of sharks (*Carcharhinus amblyrhynchos* and
*Carcharhinus melanopterus*) between fixed location autonomous arrays. In this
example, nodes represent locations within the array, *intralayer edges*
represent individual movement frequency, and *interlayer edges* represent
contact probability among individuals at a given node [@Mourier_2019]. Although
movement between fixed arrays is inherent, networks are not constructed based on
continuous measures of dyadic or collective movement [@Long_2014]. Our case
study builds on the work of Mourier et al. (2019) and explicitly integrates
continuous and simultaneous measures of animal movement through space and time
as layers of a multilayer social network.

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine
gunieafowl (*Acryllium vulturinum*), GPS movement data were used to generate
monolayer association networks within a multilayer network [@Papageorgiou_2019].
However, despite potential to generate networks in different spatial or social
contexts, layers in the multilayer network represented social associations
within and between groups. Continuous collection of animal relocation data using
GPS technology can be used to measure how groups of animals make collective
movement decisions in the context of their local environment
[@Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017] and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups [@Bode_2011]. This movement layer could be made at a
fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale [global migration:
@Flack_2018]. Our case study integrates variation in temporal and spatial scale
to assess the role of scale in predicting multilayer networks. As one might
expect, we found increasing social threshold results in saturation of potential
social connections (Figure \@ref(fig:socres)).<!--TODO: what??, while increasing temporal scale had little effect
on habitat-specific networks (Figure \@ref(fig:))-->. Another potential application would be
to generate social networks during times when animals are engaged in different
behaviours, including travelling, interacting, or foraging [@Muller_2018;
@Finn_2019] and construct a multilayer network based on behaviour-specific
monolayer networks [e.g. @Smith_Aguilar_2019].

## Habitat selection and space use
Social structure and habitat selection are intimately linked: individuals must
share space to interact and the distribution of resources influences group
dynamics and home range size [@Webber_2018]. Social structure is influenced by
resource availability and seasonality [@Peignier_2019], configuration [@He_2019]
and fragmentation [@Banks_2007]. Habitat selection, the disproportionate use of
available resources or landscape features by animals [@Manly_1993; @Boyce_1999],
is a scale-dependent process [@Mayor_2009]. Multilayer networks represent a
novel framework for considering the influence of habitat selection and space use
on sociality across scales.

<!-- TODO: point to figure 1 -->
Habitat selection depends on the scale at which it is measured and a singular
scale is insufficient for generalizing across scales [@Mayor_2009]. Habitat
selection is measured using resource selection functions (RSFs), based on the
proportion of used and available resources for an animal [@Manly_1993;
@Boyce_1999]. Johnson (1980) identified four orders of habitat selection: (1)
the geographical range of a species, (2) the home range of an individual, the
(3) habitat patches within the home range and (4) specific resources within a
habitat patch. Kie et al. [-@Kie_2002] found that, independently, third order of
selection was insufficient for capturing landscape attributes selected for by
mule deer (*Odocoileus hemionus*), at a scale larger than their home range size.
Moving past this hierarchical perspective, Mayor (2007) emphasized the
importance of considering a broad continuum of scales to reveal scale-dependent
selection and avoid the bias of predefined scales assumed to be associated with
certain behaviours. Despite animals selecting for different landscape attributes
at different scales [@Mayor_2009], the use of multiscale analyses for habitat
selection is still critically lacking [@McGarigal_2016].

<!-- TODO: weak --> 
Multilayer networks can be used to explicitly consider
habitat selection and sociality across scales. Animal social systems can be
parsed by spatial contexts, defined, for example, by habitat types as in Figure
\@ref(fig:lcres). The effect of seasonality of resources can can be directly
measured by differences in habitat selection and sociality across temporal
layers as in Figure \@ref(fig:winpos) and Figure \@ref(fig:winlen). Habitat
selection can be explicitly integrated using network layers defined by selection
coefficients for individuals. These examples of explicit integrations of habitat selection
and sociality across social, spatial and temporal scales highlight the potential
for novel insights from multilayer networks.

@@

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. We demonstrate the potential novel analyses that vary social, spatial,
and temporal scale within a multilayer framework. In our caribou case study, we
show that (1) multilayer social networks can be, but are not always,
scale-dependent, and (2) the use of multilayer networks could be useful for
identifying habitat or season-specific social processes. Our analyses of social,
spatial, and temporal scales in a multilayer network framework highlight the potential
to identify scales at which different properties of layers within a multilayer
network show the most variability. Furthermore, we discuss the potential
integration of individual phenotypes and fitness, collective movement, and
habitat selection for future analyses considering spatial ecology using
multilayer networks.

Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. In addition, recent technological
advances have lead to increased data complexity and resolution. Remotely sensed
imagery collected by modern satellites provide, for example, 10 m spatial
resolution optical imagery at a temporal revisit period of five days at the
equator [@Drusch_2012]. Similarly, animal tracking devices have become smaller
allowing their deployement on a wider variety of organisms [@Wikelski_2007] and
at a higher temporal resolution [@B_rger_2016]. Given this increased
availability of complex data, it is important to consider using a broad
continuum of scales to fully capture ecological processes, carefully select and
combine sources of data with different temporal and spatial scales, and
establish a clear link between scales of observation and meaningful biology.

@@

<!-- JWT: I'm having a hard time getting to where I want to go, but basically I want to highlight that using GPS relocation data is reflective of more traditional observed networks like they wanted when they accepted our abstract, and this provides opportunities for social networks, particularly multilayer networks. It may be easier to scale relocation data and other landscape variables and can link to smaller scale observed interactions, providing a bridge between different types of data. -->

@@

# Acknowledgements

We thank all members of the Wildlife Evolutionary Ecology Lab, including J.
Hendrix, K. Kingdon, S. Boyle, C. Prokopenko, I. Richmond, J. Hogg, and L.
Newediuk for their comments on previous versions of this manuscript as well as
D. C. Schneider for inspiration and helpful discussions about scale. Funding for
this study was provided by a Vanier Canada Graduate Scholarship to QMRW and a
NSERC Discovery Grant to EVW.

>>>>

# Discussion

<!-- TODO: edge overlap can indicate if some associations are dependent on context -->
<!-- sufficient observations is essential for capturing reliable networ -->

@@

<!-- TODO add? each ordinal network layer consisted of avg 380 fixes * ID -->
<!-- -------------------------------------------------- -->

<!-- 1: refresh why and what  -->
<!-- importance of scale in ml -->
Our results highlight the influence of social, spatial and temporal scale on multilayer networks. We found increasing connectedness between individuals with increasing

Our examinations is motivated by behavioural, landscape, and spatial ecology to
build a conceptual and analytical framework for scale-dependent multilayer
networks, and we apply this framework to a case study of caribou (*Rangifer
tarandus*) in Newfoundland, Canada [@Peignier_2019]. We use this case study to
address two principal questions associated with scale-dependent multilayer
networks. First, to what degree does the social scale of relationships, spatial
scale of landscapes, and temporal scale of analysis influence our ability to
interpret complex social systems? Second, can variation in the social, spatial,
and temporal scale improve predictions associated with biological differences in
seasonal resource availability and social association? We conclude with
perspectives for further development and application of scale in multilayer
networks with respect to phenotypes and individual fitness,
movement ecology and collective movement, and habitat selection
and space use.

<!-- 2: measuring sociality is powerful, fitness -->

Monolayer network approaches have been increasingly used to link social network
position with individual fitness, including in primates [@Brent_2017;
@Thompson_2019], dolphins [@Stanton_2012; @Stanton_2011], and birds
[@McDonald_2007; @Royle_2012]. Social scale have emerged as important for
understanding fitness consequences in the context of social network positions
<!-- TODO JWT check --> [@Almeling_2016; @Berger_2015; @Brent_2017;
@Holekamp_2011]. Defining the social scale at which to generate networks is an
essential consideration for social network analysis with direct impacts on
outcomes <!--blah-->[@Castles_2014; @Carter_2015; @Farine_2015a]. Proximity-based social networks generated using GPS location data require using a social scale, specifically a spatial distance threshold, that defines social associations. 
In 
threshold to

Multilayer networks

In contrast to multilayer networks formed by combining layers defined
by different types of interactions, multilayer networks with GPS data are limited to 
proximity based measures. An alternative which has been explored is using 
hidden markov models behavioural states, within which individuals are associating. [@Jones_2020]

<!-- types of relationships, ML nets -->
<!-- GPS caveat HMM -->

Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a].

For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017].

Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

social scale results

## Social scale

Social scale is a major consideration for social network analysis,
especially in the context of different types of relationships, such as grooming
and association, that are typically examined separately while answering a single
question [@Castles_2014; @Carter_2015; @Farine_2015a]. Furthermore, it is a
common trope that individuals must be associating to interact, and many studies
therefore assume that proximity is proxy for interacting [@Farine_2015a]. For
instance, baboons need to be in proximity to groom each other. However, animals
have certain interactions that do not require proximity such as long-distance
vocalization and olfactory signals [@Carter_2015]. With the advancement of
biologging technology, we have the ability to record social interactions that do
not require proximity, particularly long-distance communication networks with
devices like microphone arrays [@Snijders_2017]. Using relocation data, our case
study results indicate that there are different social patterns in different
habitat types, likely indicating different types of interactions. When the
social threshold is greater than approximately 50 m during the winter,
individuals demonstrate dramatically different graph strength (Figure 
\@ref(fig:socres)) in different habitat types, having higher graph strength in
forage and open habitats where they can likely see each other and still interact
with visual cues unlike in forest habitats. Varying social scales can have
different fitness repercussions for individuals. Rhesus macaques (*Macaca
mulatta*) had higher survivorship when they had fewer but stronger, more stable
social associates and grooming partners rather than simply more associates
[@Ellis_2019]. This finding highlights that different qualities of social scales
have lasting consequences. Multilayer networks provide the opportunity to
address these types of questions involving social scale with all the individual
layers of interaction integrated in a holistic way.

Social structure and habitat selection are intimately linked: individuals must
share space to interact and the distribution of resources influences group
dynamics and home range size [@Webber_2018]. Social structure is influenced by
resource availability and seasonality [@Peignier_2019], configuration [@He_2019]
and fragmentation [@Banks_2007]. Habitat selection, the disproportionate use of
available resources or landscape features by animals [@Manly_1993; @Boyce_1999],
is a scale-dependent process [@Mayor_2009]. Multilayer networks represent a
novel framework for considering the influence of habitat selection and space use
on sociality across scales.

## Spatial scale

Habitat selection depends on the scale at which it is measured and a singular
scale is insufficient for generalizing across scales [@Mayor_2009]. Habitat
selection is measured using resource selection functions (RSFs), based on the
proportion of used and available resources for an animal [@Manly_1993;
@Boyce_1999]. Johnson (1980) identified four orders of habitat selection: (1)
the geographical range of a species, (2) the home range of an individual, the
(3) habitat patches within the home range and (4) specific resources within a
habitat patch. Kie et al. [-@Kie_2002] found that, independently, third order of
selection was insufficient for capturing landscape attributes selected for by
mule deer (*Odocoileus hemionus*), at a scale larger than their home range size.
Moving past this hierarchical perspective, Mayor (2007) emphasized the
importance of considering a broad continuum of scales to reveal scale-dependent
selection and avoid the bias of predefined scales assumed to be associated with
certain behaviours. Despite animals selecting for different landscape attributes
at different scales [@Mayor_2009], the use of multiscale analyses for habitat
selection is still critically lacking [@McGarigal_2016].

Multilayer networks can be used to explicitly consider
habitat selection and sociality across scales. Animal social systems can be
parsed by spatial contexts, defined, for example, by habitat types as in Figure
\@ref(fig:lcres). The effect of seasonality of resources can can be directly
measured by differences in habitat selection and sociality across temporal
layers as in Figure \@ref(fig:winpos) and Figure \@ref(fig:winlen). Habitat
selection can be explicitly integrated using network layers defined by selection
coefficients for individuals. These examples of explicit integrations of habitat selection
and sociality across social, spatial and temporal scales highlight the potential
for novel insights from multilayer networks.

## Temporal scale
Temporal scale is also important for understanding the repercussions of social
phenotypes. Varying the time window length and position for caribou, we found
there was a lot of variation in the multilayer network metrics (Figure
\@ref(fig:winlen), Figure \@ref(fig:winpos)). These changes in the outcome
depending on how the time window is defined highlight how critical it is to
properly define the time window appropriately to address the questions and
hypotheses of interest. For instance, social phenotypes can change over ontogeny
in ways that can affect fitness [@Berger_2015; @Brent_2017]. Thus, it is
critical to carefully consider the temporal scale that is meaningful for the
proposed hypotheses. Further complicating the issue of selecting suitable time
scales across ontogeny, individuals progress through development at different
rates [@Tarka_2018]. This variation can make defining appropriate temporal
scales for testing the fitness implications of an individual’s social phenotype
over different stages of ontogeny even more difficult when they vary by
individual. However, considering fine temporal scales provide insight across
developmental stages that may otherwise be missed [@Turner_2017]. Defining
temporal scales is a challenge in multilayer networks just as in monolayer
networks; however, multilayer networks provide metrics that enable researchers
to look at similarities and differences in separate layers representing
different stages of ontogeny, or seasons as in our case study, which can be
difficult to detect.

Furthermore, multilayer networks provide a framework
to test hypotheses linking social phenotypes and fitness across temporal scales.
Because of the non-independent nature of social interactions and associations,
especially over time, it is hard to test hypotheses about social dynamics with a
suitable null control [@Farine_2017; @Proskurnikov_2017]. Using the framework of
multilayer networks, making layers of different temporal periods provides a way
to make appropriate null models to test questions about the dynamics of an
individual’s social network position over multiple aggregated time periods.

## Synth

## Etc

@@

@@

Multilayer networks provide unique opportunities and challenges, linking these
various scales, for understanding the social phenotypes and environments of
individuals and their fitness outcomes.

This multilevel network
metric could be used to assess the relationship between an individual’s social
phenotype and fitness. It would also be a meaningful reduction of many
network metrics that could help simplify statistical analyses and deal with the
frequent correlations between social network metrics in different types of
networks [@Castles_2014].

@@

<!-- fixed array -->
Although multilayer animal social networks are relatively novel [@Silk_2018],
movement behaviour within multilayer networks has been considered. Spatial
networks are physical locations (nodes) which are connected in a network when
individuals move between said locations (edges). For example, spatial networks
were applied to monitor movement of sharks (*Carcharhinus amblyrhynchos* and
*Carcharhinus melanopterus*) between fixed location autonomous arrays. In this
example, nodes represent locations within the array, *intralayer edges*
represent individual movement frequency, and *interlayer edges* represent
contact probability among individuals at a given node [@Mourier_2019]. Although
movement between fixed arrays is inherent, networks are not constructed based on
continuous measures of dyadic or collective movement [@Long_2014]. Our case
study builds on the work of Mourier et al. (2019) and explicitly integrates
continuous and simultaneous measures of animal movement through space and time
as layers of a multilayer social network.

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine
gunieafowl (*Acryllium vulturinum*), GPS movement data were used to generate
monolayer association networks within a multilayer network [@Papageorgiou_2019].
However, despite potential to generate networks in different spatial or social
contexts, layers in the multilayer network represented social associations
within and between groups. Continuous collection of animal relocation data using
GPS technology can be used to measure how groups of animals make collective
movement decisions in the context of their local environment
[@Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017] and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups [@Bode_2011]. This movement layer could be made at a
fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale [global migration:
@Flack_2018].

Our case study integrates variation in temporal and spatial scale
to assess the role of scale in predicting multilayer networks. As one might
expect, we found increasing social threshold results in saturation of potential
social connections (Figure \@ref(fig:socres)).<!--TODO: what??, while increasing temporal scale had little effect
on habitat-specific networks (Figure \@ref(fig:))-->. Another potential application would be
to generate social networks during times when animals are engaged in different
behaviours, including travelling, interacting, or foraging [@Muller_2018;
@Finn_2019] and construct a multilayer network based on behaviour-specific
monolayer networks [e.g. @Smith_Aguilar_2019].

## Habitat selection and space use

@@

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. We demonstrate the potential novel analyses that vary social, spatial,
and temporal scale within a multilayer framework. In our caribou case study, we
show that (1) multilayer social networks can be scale-dependent, and (2) the use of multilayer networks could be useful for
identifying habitat or season-specific social processes. Our analyses of social,
spatial, and temporal scales in a multilayer network framework highlight the potential
to identify scales at which different properties of layers within a multilayer
network show the most variability.

Furthermore, we discuss the potential
integration of individual phenotypes and fitness, collective movement, and
habitat selection for future analyses considering spatial ecology using
multilayer networks.

<!-- y -->
Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. In addition, recent technological
advances have lead to increased data complexity and resolution. Remotely sensed
imagery collected by modern satellites provide, for example, 10 m spatial
resolution optical imagery at a temporal revisit period of five days at the
equator [@Drusch_2012]. Similarly, animal tracking devices have become smaller
allowing their deployement on a wider variety of organisms [@Wikelski_2007] and
at a higher temporal resolution [@B_rger_2016]. Given this increased
availability of complex data, it is important to consider using a broad
continuum of scales to fully capture ecological processes, carefully select and
combine sources of data with different temporal and spatial scales, and
establish a clear link between scales of observation and meaningful biology.
<!-- y -->

@@

@@

# Acknowledgements
<!-- jbf -->
We thank all members of the Wildlife Evolutionary Ecology Lab, including J.
Hendrix, K. Kingdon, S. Boyle, C. Prokopenko, I. Richmond, J. Hogg, and L.
Newediuk for their comments on previous versions of this manuscript as well as
D. C. Schneider for inspiration and helpful discussions about scale. Funding for
this study was provided by a Vanier Canada Graduate Scholarship to QMRW and a
NSERC Discovery Grant to EVW.
